Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-03-25 parallel.tex
final checkin of weak scaling section, including some new references.

\subsubsection{Unigrid scaling \red{(Brian)}}

@@

Figure~\ref{fig.uniscale} shows \enzo\ weak scaling results for a
sequence of scaled unigrid Lyman alpha forest calculations. These
calculations include dark matter dynamics, hydrodynamics using the
piecewise parabolic method, six-species non-equilibrium chemistry and
radiative cooling, and a uniform metagalactic ultraviolet background.
In this sequence of test calculations, we perform a weak scaling test
on up to 13,824 MPI tasks on the NICS Kraken XT4 and ORNL Jaguar XT4
supercomputers\footnote{these simulations were performed in 2008,
prior to conversion of both machines to the current-generation
systems}.  In this test, each MPI task was given a $128^3$ root grid
tile (i.e., $128^3$ grid cells containing baryon quantities) and,
initially, approximately $128^3$ dark matter particles.  The number of
grid cells was constant throughout the calculation; the number of dark
matter particles varies as they are moved from subvolume to subvolume
as structure evolves.  The grid resolution was kept at a constant
comoving size of $\simeq 40$~kpc/h, and as the core count was
increased, so was the simulation volume.  On each machine, a compute
node contained a single AMD Opteron quad-core chip (2.1 Ghz on Jaguar;
2.3 Ghz on Kraken) with 2 GB/memory per core (8 GB/total per node).
Both machines used the SeaStar2 interconnect.  In the scaling study,
calculations were run with 1, 2 or 4 MPI tasks per node.  The figure
shows cell updates per second per MPI process; perfect scaling would
be a horizontal line.

@@

Broadly, excellent scaling in \enzo's unigrid
mode is seen for a variety of problems as long as each compute core is
given an adequate amount of work to do.  For cosmological simulations, this seems to be
roughly $128^3$ cells per core.  If fewer cells per core are used, the
CPU is essentially data-starved, and poor scaling is observed due to
computing units being idle while waiting for information to be
communicated from other processes (for, e.g., boundary information or
gravity solves).  Substantially larger cell counts per core  would in
principle help scaling, but are typically impractical on
most machines due to memory limits.

As a final point, we 
up to 98,304
cores has been achieved on the NICS Kraken XT5 using a hybrid-parallel
(MPI + OpenMP)
version of \enzo\.  Excellent scaling in unigrid mode is seen as long
as each core is given enough work -- somewhere around $128^3$
cells/core seems to be about right.

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth]{figures/enzo_unigrid_weak_scaling.eps}
\caption{Unigrid scaling.  ENZO weak scaling performance for a set of
  Lyman Alpha Forest cosmology sim- ulations with constant physical
  resolution per mesh point, showing cell updates per second per
  computational core plotted as a function of the number of root grid
  tiles of dimension $128^3$ (R) in each dimension. The number of MPI
  tasks is N$ = R^3$, so R$ = 16$ on this plot corresponds to a $2048^3$ computational mesh running on 4096 MPI tasks. }
\label{fig.uniscale}
\end{center}
\end{figure}

>>>>

\subsubsection{Unigrid scaling}

@@

Figure~\ref{fig.uniscale} shows \enzo\ weak scaling results for a
sequence of scaled unigrid Lyman alpha forest calculations. These
calculations include dark matter dynamics, hydrodynamics using the
piecewise parabolic method, six-species non-equilibrium chemistry and
radiative cooling, and a uniform metagalactic ultraviolet background.
In this sequence of test calculations, we perform a weak scaling test
on up to 13,824 MPI tasks on the NICS Kraken XT4 and ORNL Jaguar XT4
supercomputers\footnote{These simulations were performed in 2008,
prior to conversion of both machines to the current-generation
systems}.  In this test, each MPI task was given a $128^3$ root grid
tile (i.e., $128^3$ grid cells containing baryon quantities) and,
initially, approximately $128^3$ dark matter particles.  The number of
grid cells was constant throughout the calculation; the number of dark
matter particles varies as they are moved from subvolume to subvolume
as structure evolves.  The grid resolution was kept at a constant
comoving size of $\simeq 40$~kpc/h, and as the core count was
increased, so was the simulation volume.  On each machine, a compute
node contained a single AMD Opteron quad-core chip (2.1 Ghz on Jaguar;
2.3 Ghz on Kraken) with 2 GB/memory per core (8 GB/total per node).
Both machines used the SeaStar2 interconnect.  In the scaling study,
calculations were run with 1, 2 or 4 MPI tasks per node.  The figure
shows cell updates per second per MPI process; perfect scaling would
be a horizontal line.

@@

Broadly, excellent scaling in \enzo's unigrid mode is seen for a
variety of problems as long as each compute core is given an adequate
amount of work to do.  For cosmological simulations, this seems to be
roughly $128^3$ cells per core.  If fewer cells per core are used, the
CPU is essentially data-starved, and poor scaling is observed due to
computing units being idle while waiting for information to be
communicated from other processes (for, e.g., boundary information or
gravity solves).  Substantially larger cell counts per core would in
principle help scaling by reducing the amount of inter-process
communication needed, but larger cell counts are typically impractical
on most machines due to memory limits.

As a final point, we observe that scaling at larger core counts has
been measured, but only with an experimental hybrid-parallel (MPI +
OpenMP) version of \enzo.  Using this version, scaling comparable to
that shown in Figure~\ref{fig.uniscale} was seen on up to 98,304 cores
on the NICS Kraken XT5 (an upgraded version of the XT4 machine used
for the scaling study shown in the figure), using 2-8 OpenMP threads
per MPI process.  Hybrid parallelism has the potential to
substantially improve scaling by reducing the amount of communication
per grid tile (as described in the previous paragraph); however, the
optimal ratio of OpenMP threads per MPI task seems to vary
substantially between computational platforms and astrophysical
problems/required \enzo\ physics modules, and thus we hesitate to
provide strict guidelines here.

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth]{figures/enzo_unigrid_weak_scaling.eps}
\caption{\enzo\ weak scaling performance for a set of Lyman alpha
forest cosmology simulations with constant comoving spatial resolution
per grid cell, showing cell updates per second per computational core
plotted as a function of the number of root grid tiles of dimension
$128^3$ (R) in each dimension.  The number of MPI tasks is N$ = R^3$,
so R$ = 16$ on this plot corresponds to a $2048^3$ computational mesh
running on 4096 MPI tasks.  This plot goes from R$ = 2$ (8 MPI tasks)
to R$ = 24$ (13,824 MPI tasks) on two supercomputers -- NICS Kraken
and ORNL Jaguar when they were Cray XT4 systems -- and using 1, 2 or 4
MPI processes per node, where each compute node contained a single
quad-core AMD Opteron CPU having a speed of 2.1 GHz on Jaguar and 2.3
GHz on Kraken.}
\label{fig.uniscale}
\end{center}
\end{figure}
